Ms 1

### 2017-07-26 ms1/ms1_display.Rmd
more streamlining

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(stringr, str_replace, str_detect)
library(forcats)
import::from(tools, md5sum)
```

@@

```{r abstract}
pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

@@

est2pct = function(x) (exp(x) - 1)*100
model_ci = function(df) {
  df %>%
    filter(model=='poisson', term=='year') %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f%% (%.2f--%.2f%%)', pct_change, ci_low, ci_high))
}

@@

total_adjusted_change_str = model_abx %>%
  filter(antibiotic=='overall') %$%
  change_str

azithro_adjusted_change_str = model_abx %>%
  filter(antibiotic=='azithromycin') %$%
  change_str

levo_adjusted_change_str = model_abx %>%
  filter(antibiotic=='levofloxacin') %$%
  change_str
```

@@

```{r adjusted_changes, cache.extra=md5sum('tables/model_abx.tsv')}
x = model_abx %>%
  filter(term=='year') %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  mutate(is_overall=antibiotic=='overall') %>%
  arrange(is_overall, pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic)) %>%
  select(-is_overall)

@@

# Azithro and levo

I used the encounter data to anotate which beneficiaries had COPD or an ARTI in that year. I used these primary diagnoses:

```{r}
read_tsv('../sex-difference/sex_codes.tsv') %>%
  filter(diagnosis_type %in% c('acute_rc', 'copd')) %>%
  kable
```

Then I ran models predicting azithromycin and levofloxacin consumption using these some standard demographic variables and a narrowly-defined health status (COPD and ARTI).

## Azithromycin

`y ~ year + start_age + is_female + is_dual + is_white + region + copd + copd:year + copd:start_age + acute_rc + acute_rc:year + acute_rc:start_age`

That is, add COPD, ARTI, and the interactions of those terms with beneficiary age (in 2011) and calendar year.

```{r, cache.extra=md5sum('tables/tbl_cohort_model_azithro.tsv')}
read_tsv('tables/tbl_cohort_model_azithro.tsv') %>%
  model_ci_cpkp() %>%
  filter(term != '(Intercept)') %>%
  select(term, cpkp, ci_low, ci_high, p.value) %>%
  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
  kable
```

Azithromycin prescribing decreases with calendar year and with age, even for those who have COPD or an ARTI.

## Levofloxacin

`y ~ year + start_age + is_female + is_dual + is_white + region + acute_rc + acute_rc:year + acute_rc:start_age`

That is, add ARTI and the interactions with beneficiary age and calendar year.

```{r, cache.extra=md5sum('tables/tbl_cohort_model_levo.tsv')}
read_tsv('tables/tbl_cohort_model_levo.tsv') %>%
  model_ci_cpkp() %>%
  filter(term != '(Intercept)') %>%
  select(term, cpkp, ci_low, ci_high, p.value) %>%
  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
  kable
```

By contrast, levofloxacin prescribing increases with calendar year and with age, including for those who have and ARTI. The overall increase with time is about half the size of the increase with time for those who have an ARTI, meaning that levofloxacin, aside from being prescribed more overall, is being prescribed more especially for those with ARTIs.

Interestingly, the effect of dual-enrollment is much higher for levofloxacin than azithromycin (about triple). Why would being poorer make you more likely to take levofloxacin?

# Diagnoses

```{r}
read_tsv('tables/model_abx.tsv') %>%
  filter(term=='year') %>%
  mutate(model=case_when(.$model=='trimethoprim/sulfamethoxazole' ~ 'TMP/SMX',
                         .$model=='amoxicillin/clavulanate' ~ 'amox/clav',
                         TRUE ~ .$model)) %>%
  mutate(model=forcats::fct_reorder(model, estimate),
         cpkp=1000*estimate) %>%
  ggplot(aes(x=model, y=cpkp)) +
  geom_point() +
  geom_linerange(aes(x=model, ymin=0, ymax=cpkp)) +
  coord_flip() +
  xlab('') + ylab('Adjusted yearly change in claims per 1k bene. per yr') +
  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
  theme_minimal()
```

# Supplemental results

## Summary of data for the entire group

```{r}
read_tsv('tables/tbl_totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sem_age, mean_n_cc, sem_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1),
         mean_n_cc=round(mean_n_cc, digits=2)) %>%
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  kable
```

## Consumption trends in both groups

```{r}
dat = read_tsv('tables/tbl_claims_by_abx.tsv')

all_dat = bind_rows(
  top_abx_f(dat) %>% mutate(group='all'),
  top_abx_f(cohort_dat) %>% mutate(group='cohort')
)

all_dat %>%
  ggplot(aes(x=year, y=cpkp, color=group)) +
  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
  geom_point() + geom_line() +
  theme_minimal() +
  ggtitle('Unadjusted consumption, overall or in cohort')
```

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(stringr, str_replace, str_detect)
library(forcats)
import::from(tools, md5sum)
```

@@

```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

@@

est2pct = function(x) (exp(x) - 1)*100
model_ci = function(df) {
  df %>%
    filter(model=='poisson') %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
}

@@

total_adjusted_change_str = model_abx %>%
  filter(term=='year', antibiotic=='overall') %$%
  change_str

azithro_adjusted_change_str = model_abx %>%
  filter(term=='year', antibiotic=='azithromycin') %$%
  change_str

levo_adjusted_change_str = model_abx %>%
  filter(term=='year', antibiotic=='levofloxacin') %$%
  change_str
```

@@

```{r adjusted_changes}
x = model_abx %>%
  filter(term=='year') %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  mutate(is_overall=antibiotic=='overall') %>%
  arrange(is_overall, pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic)) %>%
  select(-is_overall)

@@

# Prescribing practice

## Methodology

Let $R_0$ be the event where the drug of interest is prescribed on some specific day in the first study period; let $R_1$ represent the same thing in the second period. Let $D_0$ be the event where the diagnosis of interest was made within some window before the day of interest; similarly $D_1$. Thus we have $P(R_0 | D_0)$, which represents, say, the probability that levo will be prescribed following a pneumonia diagnosis in 2011. The fold change in the
probability that a drug will be prescribed for a diagnosis is
$$
\frac{P(R_1 | D_1)}{P(R_0 | D_0)} = \frac{P(R_1 \cap D_1) / P(R_0 \cap D_0)}{P(D_1) / P(D_0)}.
$$

The ratio of the joint probabilities is the ratio in the number of times a qualifying prescription-diagnosis pair appears in each year. The ratio of the diagnosis-in-window probabilities is the ratio of the number of diagnoses of that type made in each year.

(There's a way to do the age adjustment here too: just replace every appearance
of $D_y$ with $D_y \cap A_{ya}$.)

```{r rxdx}
rxdx = read_tsv('tables/rxdx.tsv')

rxdx_f = function(a) {
  rxdx %>%
    filter(abx==a) %>%
    select(-abx) %>%
    arrange(desc(f_rxdx))
}
```

## Levo

Levofloxacin prescribing increased across diagnoses. Of the respiratory diagnoses, pneumonia was the one associated with the smallest fractional increase in probability of prescription.

```{r rxdx_levo}
rxdx_f('levofloxacin') %>% filter(f_rxdx > 0.03) %>% kable
```

## Azithro

Azithromycin use decreased across diagnoses. Pneumonia experienced a greater decline.

```{r rxdx_azithro}
rxdx_f('azithromycin') %>% filter(f_rxdx > 0.01) %>% kable
```

## Azithromycin & heart disease

Azithromycin prescribing for people with heart disease decreased slightly more than the people without heart disease.

```{r model_azithro}
model_azithro = read_tsv('tables/model_azithro.tsv') %>%
  filter(model=='poisson')

nohd_est = model_azithro %>% filter(term=='year') %$% estimate
nohd_se = model_azithro %>% filter(term=='year') %$% std.error
offset_est = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% estimate
offset_se = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% std.error

hd_est = nohd_est + offset_est
hd_se = sqrt(nohd_se^2 + offset_se^2)

data_frame(heart_disease=c(FALSE, TRUE),
           estimate=c(nohd_est, hd_est),
           std.error=c(nohd_se, hd_se)) %>%
  mutate(model='poisson') %>%
  model_ci() %>%
  kable
```

Compare these changes to the overall azithromycin change: `r azithro_adjusted_change_str`.

A large fraction of beneficiaries have heart disease in each year:

```{r heart_disease_table}
read_tsv('tables/counts_heartdisease.tsv') %>%
  group_by(year) %>%
  mutate(f=n/sum(n)) %>%
  ungroup %>%
  filter(heart_disease) %>%
  kable
```

## Amox/clav

Amoxicillin increased for sinusitis, its major associated diagnosis, and also for other respiratory diagnoses, except pneumonia (for which it is associated with a very small fraction of prescritions). It seems like most of the increase is attributable to sinusitis.

```{r rxdx_amoxclav}
rxdx_f('amoxicillin/clavulanate') %>% kable
```

## Keflex

Increased prescribing of cephalexin seems t obe mostly for UTIs; prescribing probabilities for skin and soft tissue infections remained nearly steady.

```{r rxdx_cephalexin}
rxdx_f('cephalexin') %>% filter(f_rxdx > 0.017) %>% kable
```
